---
title: "FNCE611 Problem Set 3"
author: "Jessica Blum, Jordan Farrer, Marc Lanthemann"
date: '2017-02-24'
output: function(...) {

          fmt <- rmarkdown::pdf_document(toc = TRUE, toc_depth = 1, number_section = TRUE, df_print = 'kable',...)
        
          fmt$knitr$knit_hooks$size = function(before, options, envir) {
            if (before) return(paste0("\n \\", options$size, "\n\n"))
            else return("\n\n \\normalsize \n")
          }
        
          return(fmt)
        }

---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = 'center', size = 'small')
options(digits = 4)
pacman::p_load(tidyverse, googlesheets, pander, ggrepel)
suppressMessages(gs_auth(token = "/home/rstudio/Dropbox/googlesheets_token.rds", verbose = FALSE))
```

# Diversification

```{r results = 'hide'}
wfm_tsla <- gs_title("FNCE611") %>%  gs_read(ws = "3.1")
```

## a

```{r}
e_wfm <- mean(wfm_tsla$wfm) * 12
e_tsla <- mean(wfm_tsla$tsla) * 12

sd_wfm <- sd(wfm_tsla$wfm) * sqrt(12)
sd_tsla <- sd(wfm_tsla$tsla) * sqrt(12)

cc <- cor(wfm_tsla$wfm, wfm_tsla$tsla)
```

|Estimate (annualized)|Value|
|:---|:----|
|Expected Return (Whole Foods)|`r e_wfm`|
|Expected Return (Tesla)|`r e_tsla`|
|Standard Deviation (Whole Foods)|`r sd_wfm`|
|Standard Deviation (Tesla)|`r sd_tsla`|
|Correlation Coefficient|`r cc`|

## b

```{r}
w_wfm = 0.2
e_p <- w_wfm*e_wfm + (1-w_wfm)*e_tsla
```


If $w_{wfm} = 0.2$, then the expected return for the portfolio is 

\begin{align}
\ E[r_p] & = \omega_{w}E[r_{w}] + (1-\omega_{w})E[r_{t}] \\
 & = `r w_wfm`(`r e_wfm`) + (1 - `r w_wfm`)(`r e_tsla`) \\
 & = `r e_p`
\end {align}

## c

```{r echo = FALSE}
var_p <- w_wfm^2 * sd_wfm^2 + (1-w_wfm)^2 * sd_tsla^2 + 2*w_wfm*(1 - w_wfm)*cc*sd_wfm*sd_tsla
```


\begin{align}
\ Var[r_p] & = \omega_{w}^2\sigma_{w}^2 + \omega_{t}^2\sigma_{t}^2 + 2\omega_{w}\omega_{t}\rho_{w,t}\sigma_{w}\sigma_{t} \\
 & = `r w_wfm`^2\cdot `r sd_wfm`^2 + (1-`r w_wfm`)^2\cdot `r sd_tsla`^2 + 2\cdot`r w_wfm`\cdot(1 - `r w_wfm`)\cdot `r cc` \cdot `r sd_wfm` \cdot `r sd_tsla` \\
 & = `r var_p`
\end {align}

So the standard deviation of the portfolio is **$\sqrt{Var[r_p]} = `r sqrt(var_p)`$**.

## d


```{r}
w_wfm_opt <- uniroot(function(w) w*e_wfm + (1-w)*e_tsla -.2, interval=c(0,1), tol=0.0001)$root
```

The weights of a portfolio showing an expected return of 20% would be **$\omega_{w} = `r w_wfm_opt`$** and  **$\omega_{t} = `r (1-w_wfm_opt)`$**.

## e

```{r}
fn_mvp <- function(par) {
  w <- par[1]
  w^2 * sd_wfm^2 + (1-w)^2 * sd_tsla^2 + 2*w*(1 - w)*cc*sd_wfm*sd_tsla
}

w_wfm_mvp <- nlminb(.5, fn_mvp, lower = 0, upper = 1)$par[1]
```

The minimum variance portfolio (MVP) would have weights **$\omega_w = `r w_wfm_mvp`$** and **$\omega_t = `r (1- w_wfm_mvp)`$**.

```{r}
data_frame(w_wfm = seq(0, 1, by = 0.01)) %>%
  mutate(
    expected_std = sqrt(w_wfm^2 * sd_wfm^2 + (1-w_wfm)^2 * 
                          sd_tsla^2 + 2*w_wfm*(1 - w_wfm)*cc*sd_wfm*sd_tsla)
    , expected_return = w_wfm*e_wfm + (1-w_wfm)*e_tsla
  ) %>%
  ggplot(aes(x = expected_std, y = expected_return, colour = w_wfm)) +
  geom_path() +
  labs(x = expression(sigma[p]), y = expression(E~group("[",X[p], "]")), 
       colour = expression(omega[w])) +
  geom_point(data = . %>% top_n(-1, expected_std),  
             aes(x = expected_std, y = expected_return), colour = 'red') +
  scale_x_continuous(limits = c(0, NA))
```

## f

```{r}
fn_rfr <- function(par) {
  w <- par[1]
  sharpe <- ((w*e_wfm + (1-w)*e_tsla) - 0.02) / sqrt(w^2 * sd_wfm^2 + (1-w)^2 * sd_tsla^2 + 
                                                       2*w*(1 - w)*cc*sd_wfm*sd_tsla)
  return(-sharpe)
}
max_w <- nlminb(.5, fn_rfr, lower = 0, upper = 1)$par[1]
max_sharpe_ratio <- ((max_w*e_wfm + (1-max_w)*e_tsla) - 0.02) / sqrt(max_w^2 * sd_wfm^2 +
                      (1-max_w)^2 * sd_tsla^2 + 2*max_w*(1 - max_w)*cc*sd_wfm*sd_tsla)
```

A portfolio with the highest Sharpe ratio would have weights **$\omega_w = `r max_w`$** and **$\omega_t = `r (1- max_w)`$**.

```{r}
efficient_mix <- 
  data_frame(expected_std = 
         sqrt(max_w^2 * sd_wfm^2 + (1-max_w)^2 * sd_tsla^2 + 
                2*max_w*(1 - max_w)*cc*sd_wfm*sd_tsla)
       , expected_return = max_w*e_wfm + (1-max_w)*e_tsla
  )

data_frame(w_wfm = seq(0, 1, by = 0.01)) %>%
  mutate(
    expected_std = sqrt(w_wfm^2 * sd_wfm^2 + (1-w_wfm)^2 * sd_tsla^2 + 
                          2*w_wfm*(1 - w_wfm)*cc*sd_wfm*sd_tsla)
    , expected_return = w_wfm*e_wfm + (1-w_wfm)*e_tsla
  ) %>%
  ggplot(aes(x = expected_std, y = expected_return, colour = w_wfm)) +
  geom_path() +
  labs(x = expression(sigma[p]), y = expression(E~group("[",X[p], "]")),
       colour = expression(omega[w])) +
  geom_point(data = . %>% top_n(-1, expected_std),  
             aes(x = expected_std, y = expected_return), colour = 'red') +
  scale_x_continuous(limits = c(0, NA), expand = c(0,0)) +
  scale_y_continuous(limits = c(0, NA)) +
  geom_abline(aes(slope = max_sharpe_ratio, intercept = 0.02), size = .2) +
  geom_point(data = efficient_mix, aes(x = expected_std, y = expected_return),
             colour = 'dark green')
```

## g

\begin{align}
\ E[r_p] & = \omega_{riskless\;asset}E[r_{riskless\,asset}] + (1- \omega_{riskless\;asset})E[r_{mix}] \\
  0.25 & = \omega_{riskless\;asset}0.02 + (1- \omega_{riskless\;asset})`r efficient_mix$expected_return[1]`
\end{align}

```{r}
w_investor_rf <- uniroot(function(w) 0.02*w + (1-w)*efficient_mix$expected_return[1] - 0.25, 
                         interval=c(0,1), tol=0.0001)$root
```

For a portfolio with an expected return of 25%, **$\omega_{riskless\;asset} = `r w_investor_rf`$** and **$\omega_{mix} = `r 1 - w_investor_rf`$**. Thus, **$\omega_w = `r max_w * (1 - w_investor_rf)`$** and **$\omega_t = `r (1- max_w) * (1 - w_investor_rf)`$**.

# Efficient Portfolios, Capital Market Line

## a

```{r}
question2 <- 
  data_frame(
    label = LETTERS[1:8]
    , e_r = c(.1, .125, .15, .16, .17, .18, .18, .2)
    , sd_r = c(.23, .21, .25, .29, .29, .32, .35, .45)
  )
  
question2 %>%  
  ggplot(aes(x = sd_r, y = e_r)) + 
  geom_point() +
  labs(x = expression(sigma[r]), y = expression(E~group("[",r, "]")), 
       colour = expression(omega[w])) +
  ggrepel::geom_text_repel(aes(x = sd_r, y = e_r, label = label)) +
  scale_y_continuous(labels = scales::percent) +
  scale_x_continuous(labels = scales::percent)
```

## b

Portfolios A and D are inefficient. A is inefficient because for less risk you can get more return with portfolio B. D is inefficient because for the same risk you can get a greater return with E.

## c 

We see from the table below, that if you can borrow and lend at an interest rate of 12%, **portfolio F** has the highest Sharpe Ratio.

```{r}
question2 %>%
  mutate(
    sharpe_ratio = (e_r - 0.12) / sd_r
  ) %>%
  arrange(desc(sharpe_ratio)) %>%
  pander()
```

## d

```{r}
q4 <- 
  question2 %>%
    mutate(
      w = .25 / sd_r
      , r_p = w*e_r +(1-e_r)*0.12
      ) %>%
    arrange(desc(r_p))

q4 %>% pander()
```

If you are willing to tolerate a standard deviation of 25% you would borrow `r 100 * (1 - q4$w[1])`% of your investment at the 12% interest rate and invest `r 100 * q4$w[1]`% of your investment in portfolio B.

# CAPM, Security Market Line (1)

+ By definition we know that the Beta of the market portfolio M must be 1. 
+ We find the $sd(X) = \sqrt{\frac{\rho}{\beta}} = \sqrt{\frac{.4}{1.2}}$
+ We use the fact that $r - r_f = \beta(r_m - r_f)$ to find $r_x = 9$ and $r_Q = 13$

```{r}
data_frame(
  `Title` = c("Stock X", "Efficient Portfolio Q","Market Portfolio M","Riskless Asset")
  , `Expected Return` = c("9%", "13%", "8%", "3%")
  , Beta = c(1.2, 2.0, 1, 0.0)
  , `Standard Deviation` = c("57.74%", NA, "20%", "0%")
) %>%
  pander(missing = "")
```




# CAPM, Security Market Line (2)
